Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices

### 2025-03-01 jso.qmd
caption on fig 1

```{r}
#| label: fig-example-functions
#| echo: false
#| fig-width: 9
#| fig-height: 6
#| out-width: 80%
#| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection. A Huber plot shows the PP index values for all 1D data projections in polar coordinates."
#proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
best_proj <- matrix(c(-0.44157005,  0.80959831,
                      0.14622089, -0.08228193,
                      -0.08454314, -0.23234287,
                      0.70367852,  0.28185021,
                      0.27876825,  0.44819691,
                      0.45123453,  0.05896646), ncol=2, byrow=T)

@@

```{r}
#| label: fit-logistic
sim_df2 <- sim_df |>   
  mutate(n_jellies = n_jellies/10, max_tries = max_tries/10) |> 
  rename(proportion = P_J)

mod1 <- glm(proportion ~ smoothness + squintability + d + n_jellies + max_tries, data = sim_df2, family = binomial) 
```

@@

```{r}
#| tbl-pos: H
#| label: tbl-joint-output
library(kableExtra)
mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = binomial)
mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = binomial)
summary_table <- bind_rows(
  glance(mod2) %>% mutate(Model = "Squintability"),
  glance(mod3) %>% mutate(Model = "Smoothness")
) %>%
  select(Model, null.deviance, df.null, logLik, df.residual, AIC, BIC) %>%
  rename(
    `Null Deviance` = null.deviance,
    `DF Null` = df.null,
    `Log-Likelihood` = logLik,
    `DF Residual` = df.residual
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))
summary_table %>%
  kable(caption = "Model summaries for squintability and smoothness separately") %>%
  kable_styling(font_size = 10) %>%
  column_spec(1, bold = TRUE, width = "5em")
```

>>>>

```{r}
#| label: fig-example-functions
#| echo: false
#| fig-width: 9
#| fig-height: 6
#| out-width: 80%
#| fig-cap: "Examples of PP indexes with large (top row) and small (bottom row) squint angles, shown with a Huber plot, and histogram of the projected data corresponding to the optimal projection. A Huber plot shows the PP index values for all 1D data projections in polar coordinates. The skewness index on the trimodal data is also smoother than the binned normality index on RANDU."
#proj <- animate_xy(flea[,1:6], guided_tour(lda_pp(flea$species)))
best_proj <- matrix(c(-0.44157005,  0.80959831,
                      0.14622089, -0.08228193,
                      -0.08454314, -0.23234287,
                      0.70367852,  0.28185021,
                      0.27876825,  0.44819691,
                      0.45123453,  0.05896646), ncol=2, byrow=T)

@@

```{r}
#| label: fit-logistic
sim_df2 <- sim_df |>   
  mutate(n_jellies = n_jellies/10, max_tries = max_tries/10) |> 
  rename(proportion = P_J) #|>
  #filter(d < 10)

mod1 <- glm(proportion ~ smoothness + squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial) 
mod2 <- glm(proportion ~ squintability + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
mod3 <- glm(proportion ~ smoothness + d + n_jellies + max_tries, data = sim_df2, family = quasibinomial)
mod4 <- glm(proportion ~ squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
mod5 <- glm(proportion ~ smoothness*d + squintability*d + n_jellies, data = sim_df2, family = quasibinomial)
```

@@

```{r}
#| tbl-pos: H
#| label: tbl-joint-output
library(kableExtra)
summary_table <- #bind_rows(
  glance(mod1) |> #%>% mutate(Model = "Full"),
  #glance(mod2) %>% mutate(Model = "Squintability"),
  #glance(mod3) %>% mutate(Model = "Smoothness")
#  glance(mod4) %>% mutate(Model = "Interaction")
#) |>
  select(#Model, 
    deviance, df.residual, null.deviance, df.null, #logLik, 
         ) |> #, AIC, BIC) %>%
  rename(
    `Null Deviance` = null.deviance,
    `DF Null` = df.null,
    `Deviance` = deviance,
    `DF Residual` = df.residual#,
    #`Log-Likelihood` = logLik,
  ) |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
summary_table |>
  kable(caption = "Fit statistics for the model.") |>
  kable_styling(font_size = 10) |>
  row_spec(0, bold = TRUE) #, width = "5em")
```

### 2025-03-01 jso.tex
caption on fig 1

\caption{\label{fig-example-functions}Examples of PP indexes with large
(top row) and small (bottom row) squint angles, shown with a Huber plot,
and histogram of the projected data corresponding to the optimal
projection. A Huber plot shows the PP index values for all 1D data
projections in polar coordinates.}

@@

\begin{tcolorbox}[enhanced jigsaw, colbacktitle=quarto-callout-note-color!10!white, left=2mm, toptitle=1mm, colback=white, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, leftrule=.75mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, breakable, opacityback=0, titlerule=0mm, bottomtitle=1mm, rightrule=.15mm, arc=.35mm, toprule=.15mm, opacitybacktitle=0.6, coltitle=black]

@@

\toprule
\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
\midrule\addlinespace[2.5pt]
Smoothness & 1.00 (0.81 to 1.20) & 0.961 \\ 
Squintability & 1.24 (1.05 to 1.58) & \textbf{0.033} \\ 
Dimension & 0.57 (0.33 to 0.87) & \textbf{0.022} \\ 
Number of jellyfish & 1.32 (1.06 to 1.71) & \textbf{0.022} \\ 
Maximum number of tries & 1.06 (0.85 to 1.35) & 0.603 \\ 
\bottomrule

@@

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{5em}rrrrrr}

\caption{\label{tbl-joint-output}Model summaries for squintability and
smoothness separately}

@@

\\
\toprule
Model & Null Deviance & DF Null & Log-Likelihood & DF Residual & AIC & BIC\\
\midrule
\textbf{Squintability} & 49.29 & 75 & -26.44 & 71 & 62.87 & 74.53\\
\textbf{Smoothness} & 49.29 & 75 & -33.23 & 71 & 76.47 & 88.12\\
\bottomrule

>>>>

\caption{\label{fig-example-functions}Examples of PP indexes with large
(top row) and small (bottom row) squint angles, shown with a Huber plot,
and histogram of the projected data corresponding to the optimal
projection. A Huber plot shows the PP index values for all 1D data
projections in polar coordinates. The skewness index on the trimodal
data is also smoother than the binned normality index on RANDU.}

@@

\begin{tcolorbox}[enhanced jigsaw, breakable, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, opacityback=0, bottomrule=.15mm, opacitybacktitle=0.6, rightrule=.15mm, colback=white, leftrule=.75mm, bottomtitle=1mm, toptitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, coltitle=black]

@@

\toprule
\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
\midrule\addlinespace[2.5pt]
Smoothness & 1.00 (0.90 to 1.11) & 0.928 \\ 
Squintability & 1.24 (1.13 to 1.39) & \textbf{<0.001} \\ 
Dimension & 0.57 (0.44 to 0.73) & \textbf{<0.001} \\ 
Number of jellyfish & 1.32 (1.17 to 1.50) & \textbf{<0.001} \\ 
Maximum number of tries & 1.06 (0.94 to 1.21) & 0.333 \\ 
\bottomrule

@@

\begin{longtable}[t]{rrrr}

\caption{\label{tbl-joint-output}Fit statistics for the model.}

@@

\\
\toprule
\textbf{Deviance} & \textbf{DF Residual} & \textbf{Null Deviance} & \textbf{DF Null}\\
\midrule
20.69 & 70 & 49.29 & 75\\
\bottomrule
